---
title: "dataset_cleaning"
date: "2022-12-01"
output: github_document
---
```{r}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
library(viridis)
library(fastDummies)
set.seed(1)
```

Import original dataset

```{r}
childcare_inspection_df = read_csv("./data/DOHMH_Childcare_Center_Inspections.csv") %>% 
janitor::clean_names()
```

Basic data cleaning

1. We select 22 key variables in this dataset to finish our analysis

2. We separate the "inspection_summary_result" variable into two new variables: "inspection_summary" and

"inspection_result" to better show the summary and the result of the inspection for each center

3. Drop NA

4. Create a new variable "educational_worker_ratio"

5. Make all data in "program_type" and "facility_type" columns show in the same format : lower case

```{r}
childcare_inspection_df = childcare_inspection_df %>% 
  select(center_name, borough, zip_code, status, age_range, maximum_capacity,program_type, facility_type, 
         child_care_type, violation_category,
         violation_status,violation_rate_percent:average_critical_violation_rate,regulation_summary,
         inspection_summary_result) %>%
  separate(inspection_summary_result,into = c("inspection_summary","inspection_result"),sep = "-") %>% 
  drop_na(zip_code, age_range, violation_rate_percent,public_health_hazard_violation_rate, critical_violation_rate) %>% 
  filter(maximum_capacity != 0) %>% 
  mutate(
    educational_worker_ratio = total_educational_workers/maximum_capacity,
    program_type = tolower(program_type),
    facility_type = tolower(facility_type)
  )
```
6. Create dummy variables into a new dataframe for regression

```{r}
childcare_reg_df = childcare_inspection_df %>% 
  dummy_cols(select_columns = "borough", remove_first_dummy = TRUE) %>%
  dummy_cols(select_columns = "status", remove_first_dummy = TRUE) %>% 
  dummy_cols(select_columns = "program_type", remove_first_dummy = TRUE) %>% 
  dummy_cols(select_columns = "facility_type", remove_first_dummy = TRUE) %>%
  dummy_cols(select_columns = "child_care_type", remove_first_dummy = TRUE) %>% 
  janitor::clean_names() %>% 
  select(-borough, -status, -program_type, -facility_type, -child_care_type)
```
