---
title: "dataset_cleaning"
date: "2022-12-01"
output: github_document
---
```{r}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
library(viridis)
library(fastDummies)
set.seed(1)
```

1. Import original dataset

2. Remove repeated data

```{r}
childcare_inspection_df = read_csv("./data/DOHMH_Childcare_Center_Inspections.csv") %>% 
janitor::clean_names() %>% 
distinct()
```

Basic data cleaning

1. We select 22 key variables in this dataset to finish our analysis

2. Drop NA

3. Create a new variable "educational_worker_ratio"

4. Make all data in "program_type" and "facility_type" columns show in the same format : lower case

```{r}
childcare_inspection_df = childcare_inspection_df %>% 
  select(center_name, borough, zip_code, status, age_range, maximum_capacity,program_type, facility_type, 
         child_care_type, violation_category,
         violation_status,violation_rate_percent:average_critical_violation_rate,regulation_summary,
         inspection_summary_result) %>%
  drop_na(zip_code, age_range, violation_rate_percent,public_health_hazard_violation_rate, critical_violation_rate) %>% 
  filter(maximum_capacity != 0) %>% 
  mutate(
    educational_worker_ratio = total_educational_workers/maximum_capacity,
    program_type = tolower(program_type),
    facility_type = tolower(facility_type),
    borough =  as.factor(borough),
    status = as.factor(status),
    program_type = as.factor(program_type),
    facility_type = as.factor(facility_type),
    child_care_type = as.factor(child_care_type),
    age_range = as.factor(age_range)
  ) %>% 
  filter(program_type != "school age camp")
  write_csv(childcare_inspection_df,file = "childcare_inspection.csv",na = "NA", append = FALSE)
```


We calculated a new violation rate for each distinct program using violation category column.
```{r}
center_specific_df = childcare_inspection_df %>% 
  relocate(center_name, program_type) %>% 
  group_by(center_name, program_type) %>% 
  mutate(
    n_na = sum(is.na(violation_category)), 
    n_violation = sum(!is.na(violation_category)), 
    rate = n_violation/(n_violation + n_na)) %>% 
  arrange(center_name, program_type)
  write_csv(center_specific_df,file = "center_specific.csv",na = "NA", append = FALSE)
```






