---
title: "linear"
date: "2022-12-08"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
library(viridis)
library(fastDummies)
set.seed(1)
```

1. Import original dataset

2. Remove repeated data

```{r}
childcare_inspection_df = read_csv("./data/DOHMH_Childcare_Center_Inspections.csv") %>% 
janitor::clean_names() %>% 
distinct()
```

Basic data cleaning

1. We select 22 key variables in this dataset to finish our analysis

2. Drop NA

3. Create a new variable "educational_worker_ratio"

4. Make all data in "program_type" and "facility_type" columns show in the same format : lower case

```{r}
childcare_inspection_df = childcare_inspection_df %>% 
  select(center_name, borough, zip_code, status, age_range, maximum_capacity,program_type, facility_type, 
         child_care_type, violation_category,
         violation_status,violation_rate_percent:average_critical_violation_rate,regulation_summary,
         inspection_summary_result) %>%
  drop_na(zip_code, age_range, violation_rate_percent,public_health_hazard_violation_rate, critical_violation_rate) %>% 
  filter(maximum_capacity != 0) %>% 
  mutate(
    educational_worker_ratio = total_educational_workers/maximum_capacity,
    program_type = tolower(program_type),
    facility_type = tolower(facility_type),
    borough =  as.factor(borough),
    status = as.factor(status),
    program_type = as.factor(program_type),
    facility_type = as.factor(facility_type),
    child_care_type = as.factor(child_care_type),
    age_range = as.factor(age_range)
  ) 
```


We calculated a new violation rate for each distinct program using violation category column.
```{r}
center_specific_df = childcare_inspection_df %>% 
  relocate(center_name, program_type) %>% 
  group_by(center_name, program_type) %>% 
  mutate(
    n_na = sum(is.na(violation_category)), 
    n_violation = sum(!is.na(violation_category)), 
    rate = n_violation/(n_violation + n_na)) %>% 
  arrange(center_name, program_type)
```


Select all possible independent variables to fit a full model based on center-specific distinct records:

```{r}
# select distinct records of each center

center_distinct_df = center_specific_df %>%
  select(center_name, program_type, borough, status, maximum_capacity, total_educational_workers, educational_worker_ratio, rate) %>%
  distinct()

# fit the full model
full_lm = center_distinct_df %>%
 lm(rate ~ program_type + borough + status + educational_worker_ratio, data = .) 
summary(full_lm)

full_lm %>% 
  broom::tidy() %>%
  knitr::kable(digit = 3)

# model diagnostics: Residuals vs Fitted plot, QQ plot, Scale_location, Residuals vs Leverage
plot(full_lm)
```

Fit an alternative model:

```{r}
# fit an alternative model
alt_lm = center_distinct_df %>%
 lm(rate ~ program_type + borough + status + maximum_capacity * total_educational_workers, data = .) 
summary(alt_lm)

alt_lm %>% 
  broom::tidy() %>%
  knitr::kable(digit = 3)

# model diagnostics: residuals vs fitted plot, QQ plot, Scale_location, Residuals vs Leverage
plot(alt_lm)
```

Decision: Keep full_lm model.

### Assessing Multicollinearity

```{r}
# vif
library(performance) 
check_collinearity(full_lm)
```

This full_lm model shows low collinearity.

### Transformation

```{r}
# Boxblot to check the initial distribution of outcome and key predictor
center_distinct_df  %>%
par(mfrow = c(1,2))
boxplot(center_distinct_df$educational_worker_ratio, main = "Edu-worker Ratio")
boxplot(center_distinct_df$rate, main = "Violation Rate")
```

```{r}
# Square root transformation
full_lm_sqrt = center_distinct_df %>%
  lm(sqrt(rate) ~ program_type + borough + status + educational_worker_ratio, data = .)
summary(full_lm_sqrt)
```
Decision: Since there are 0 value in the outcome variable, log-transformation is not applicable. Moreover, sqrt transformation didn't improve model goodness of fit. Thus, we will keep the initial model and decide no transformation needed. 


### remove influential points

```{r}
# exclude two influential records

out_df = center_distinct_df[-c(381,551),]


# fit model without influential points
full_lm_out = lm(rate ~ program_type + borough + status + educational_worker_ratio, data = out_df)
summary(full_lm_out)

full_lm_out %>% 
  broom::tidy() %>%
  knitr::kable(digit = 3)
```
Decision: Influential points should be removed. 

### model validation

```{r}
library(caret)
# Use 5-fold validation and create the training sets
train = trainControl(method = "cv", number = 5)

model_caret = train(rate ~ program_type + borough + status + educational_worker_ratio,
                   data = out_df,
                   trControl = train,
                   method = 'lm',
                   na.action = na.pass)

model_caret$finalModel
print(model_caret)
```

Final linear model: full_lm_out








