---
title: "Linear regression model of center/program specific rate"
author: "Qing Zhou"
date: "2022-12-08"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
library(viridis)
library(fastDummies)
set.seed(1)
```

1. Import original dataset

2. Remove repeated data

```{r}
childcare_inspection_df = read_csv("./data/DOHMH_Childcare_Center_Inspections.csv") %>% 
janitor::clean_names() %>% 
distinct()
```

Basic data cleaning

1. We select 22 key variables in this dataset to finish our analysis

2. Drop NA

3. Create a new variable "educational_worker_ratio"

4. Make all data in "program_type" and "facility_type" columns show in the same format : lower case

```{r}
childcare_inspection_df = childcare_inspection_df %>% 
  select(center_name, borough, zip_code, status, age_range, maximum_capacity,program_type, facility_type, 
         child_care_type, violation_category,
         violation_status,violation_rate_percent:average_critical_violation_rate,regulation_summary,
         inspection_summary_result) %>%
  drop_na(zip_code, age_range, violation_rate_percent,public_health_hazard_violation_rate, critical_violation_rate) %>% 
  filter(maximum_capacity != 0) %>% 
  mutate(
    educational_worker_ratio = total_educational_workers/maximum_capacity,
    program_type = tolower(program_type),
    facility_type = tolower(facility_type),
    borough =  as.factor(borough),
    status = as.factor(status),
    program_type = as.factor(program_type),
    facility_type = as.factor(facility_type),
    child_care_type = as.factor(child_care_type),
    age_range = as.factor(age_range)
  ) 
```


We calculated a new violation rate for each distinct program using violation category column.
```{r}
center_specific_df = childcare_inspection_df %>% 
  relocate(center_name, program_type) %>% 
  group_by(center_name, program_type) %>% 
  mutate(
    n_na = sum(is.na(violation_category)), 
    n_violation = sum(!is.na(violation_category)), 
    rate = n_violation/(n_violation + n_na)) %>% 
  arrange(center_name, program_type)
```


1. Select all possible independent variables to fit a full model based on center-specific distinct records
```{r}
# select distinct records of each center

center_distinct_df = center_specific_df %>%
  select(center_name, program_type, borough, status, age_range, facility_type, educational_worker_ratio, rate) %>%
  distinct()

# fit the full model
full_lm = center_distinct_df %>%
 lm(rate ~ program_type + borough + status + age_range + facility_type + educational_worker_ratio, data = .) 
summary(full_lm)
```

#### Summary of the current data set
```{r}
skimr::skim(center_distinct_df) %>% 
  select(-c(factor.ordered, factor.n_unique, factor.top_counts,numeric.hist)) %>% 
  knitr::kable()
```


