---
title: "Statistical analysis"
author: "Qing Zhou"
date: "2022-12-09"
output: github_document
---


```{r}
library(tidyverse)
library(dplyr)
library(rvest)
library(purrr)
library(ggplot2)
library(modelr)
library(mgcv)
library(patchwork)
library(viridis)
library(fastDummies)
set.seed(1)
```

1. Import original dataset

2. Remove repeated data

```{r}
childcare_inspection_df = read_csv("./data/DOHMH_Childcare_Center_Inspections.csv") %>% 
janitor::clean_names() %>% 
distinct()
```

Basic data cleaning

1. We select 22 key variables in this dataset to finish our analysis

2. Drop NA

3. Create a new variable "educational_worker_ratio"

4. Make all data in "program_type" and "facility_type" columns show in the same format : lower case

```{r}
childcare_inspection_df = childcare_inspection_df %>% 
  select(center_name, borough, zip_code, status, age_range, maximum_capacity,program_type, facility_type, 
         child_care_type, violation_category,
         violation_status,violation_rate_percent:average_critical_violation_rate,regulation_summary,
         inspection_summary_result) %>%
  drop_na(zip_code, age_range, violation_rate_percent,public_health_hazard_violation_rate, critical_violation_rate) %>% 
  filter(maximum_capacity != 0) %>% 
  mutate(
    educational_worker_ratio = total_educational_workers/maximum_capacity,
    program_type = tolower(program_type),
    facility_type = tolower(facility_type),
    borough =  as.factor(borough),
    status = as.factor(status),
    program_type = as.factor(program_type),
    facility_type = as.factor(facility_type),
    child_care_type = as.factor(child_care_type),
    age_range = as.factor(age_range)
  ) 
```


We calculated a new violation rate for each distinct program using violation category column.
```{r}
center_specific_df = childcare_inspection_df %>% 
  relocate(center_name, program_type) %>% 
  group_by(center_name, program_type) %>% 
  mutate(
    n_na = sum(is.na(violation_category)), 
    n_violation = sum(!is.na(violation_category)), 
    rate = n_violation/(n_violation + n_na)) %>% 
  arrange(center_name, program_type)
```

### Statistical analysis

#### 1. Anova test

H0: The average number of violation in each center are not different across boroughs

H1: The average number of violation in each center are not different across boroughs
```{r}
# select distinct records of each center

center_distinct_df_2 = center_specific_df %>%
  select(center_name, program_type, borough, status, maximum_capacity, total_educational_workers, educational_worker_ratio, n_violation, rate) %>%
  distinct()

vio_borough = center_distinct_df_2 %>%
  group_by(borough, center_name) %>%
  summarize(n_violate = sum(n_violation))


fit_borough = lm(n_violate ~ borough, data = vio_borough)
summary(fit_borough)
anova(fit_borough) %>% knitr::kable(caption = "One way anova of violation frequency and borough")
```
The p-value of the above Anova test is 3.106e-10. Thus, at 0.05 significance level, we reject the null hypothesis and conclude that during the past 3 years, the average number of violations per center varies across boroughs. 

#### 2. Chi-square Test - 3 Violation Types and Boroughs

H0:  the proportions of ciritical violation among borough levels are equal;
AND  the proportions of general violation among borough levels are equal;
AND  the proportions of public health hazard violation among borough levels are equal;
vs.
H1:  not all proportions are equal.
```{r}
three_violation = 
  childcare_inspection_df %>%
  select(borough, violation_category) %>% 
  filter(violation_category != "NA")  %>% 
  count(borough, violation_category,) %>% 
  pivot_wider(
    names_from = "violation_category",
    values_from = "n"
  ) %>% 
  data.matrix() %>% 
  subset(select = -c(borough))

rownames(three_violation) <- c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")

three_violation %>% 
  knitr::kable(caption = "Results Table")
```
```{r}
chisq.test(three_violation)
```

The result of chi-square shows that , at significant level α=0.05, so we reject the null hypothesis and conclude that there does exist at least one borough’s proportion of violation amounts is different from others.
At 0.05 significance level, χ2>χcritical, with p-value < 2.2e-16. Thus, we reject the null hypothesis and conclude that there
is sufficient evidence that the category of violation is significantly different by borough.